Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-05 introduction.tex
NS and BH now small caps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@941 ccdd05d5-56c1-461a-bf98-85f6876523b4

The coalescence of compact binary systems consisting of neutron stars (NS)
and/or black holes (BH) is the most promising source of gravitational radiation
for Advanced \LIGO, Virgo, \GEO\ and \LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb,
LCGTWeb}.  Tens of binary coalescence events are expected to be observed in the
advanced detector era later this decade~\cite{Abadie:2010p10836}.

@@

To this end, we have the ambition of reporting gravitational-wave candidates
not minutes \emph{after} the merger, as has already been accomplished, but
seconds \emph{before}.  By
looking for threshold crossings before the gravitational-wave signal leaves
the detection band, it is possible to trade some signal to noise ratio (\SNR{})
for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
for NS--NS binaries in Advanced \LIGO\ assuming the event rate predictions
in~\cite{Abadie:2010p10836}.
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
be detectable $t$ seconds before coalescence.  The heavy solid line is the most
likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
light solid lines represent the confidence interval $[\dot N_{\mathrm{low}}, \dot 
N_{\mathrm{high}}]$ described in
\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in 
\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for
detection and that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a
detector in the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be 
detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be detectable 
within 100 seconds of merger.  The `zero detuning, lower power' and `\textsc{bbh} optimized' configurations are still more conducive to advance detection
but at the cost of fewer total detectable events above an \SNR\ of 8.}
\end{center}
\end{figure}

>>>>

The coalescence of compact binary systems consisting of neutron stars (\textsc{ns})
and/or black holes (\textsc{bh}) is the most promising source of gravitational radiation
for Advanced \LIGO, Virgo, \GEO\ and \LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb,
LCGTWeb}.  Tens of binary coalescence events are expected to be observed in the
advanced detector era later this decade~\cite{Abadie:2010p10836}.

@@

To this end, we have the ambition of reporting gravitational-wave candidates
not minutes \emph{after} the merger, as has already been accomplished, but
seconds \emph{before}.  By
looking for threshold crossings before the gravitational-wave signal leaves
the detection band, it is possible to trade some signal to noise ratio (\SNR{})
for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming the event rate predictions
in~\cite{Abadie:2010p10836}.
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns} sources that will
be detectable $t$ seconds before coalescence.  The heavy solid line is the most
likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
light solid lines represent the confidence interval $[\dot N_{\mathrm{low}}, \dot 
N_{\mathrm{high}}]$ described in
\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in 
\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for
detection and that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a
detector in the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be 
detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be detectable 
within 100 seconds of merger.  The `zero detuning, lower power' and `\textsc{bbh} optimized' configurations are still more conducive to advance detection
but at the cost of fewer total detectable events above an \SNR\ of 8.}
\end{center}
\end{figure}
